Contextuality and Nonlocality in \\ Discrete Quantum Theory

### 2015-11-17 DQC3.tex
DQC3.tex compiled in iopart class

\documentclass[english,12pt]{article}
\usepackage{mathrsfs}
\usepackage{fullpage}
% \documentclass[english,prl,aps,showpacs,preprint,floatfix,superscriptaddress]{revtex4-1} 
%\usepackage[T1]{fontenc}
% \usepackage[latin9]{inputenc}
% \setcounter{secnumdepth}{3}
\usepackage{url}
\usepackage{times}

\usepackage{amsthm}
\usepackage{amsmath}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% these went missing?
% \newcommand{\bra}[1]{\left\langle{#1}\right|}
% \newcommand{\ket}[1]{\left|{#1}\right\rangle}
% \newcommand{\mathscr}[1]{{\cal #1}}
%\renewcommand{\author}[1]{\bf #1 \/}
\newcommand{\affiliation}[1]{\parbox{5.5in}{\small{\it #1 \/}}}
\newcommand{\pacs}[1]{\bf #1 \/}

@@

\newcommand{\ComplexI}{\text{i}}
\allowdisplaybreaks
\newcommand{\fnorm}[1]{\mathsf{N}(#1)}

@@

%\date{\today}

@@

Our fundamental divergence from CQT is to replace continuous complex
amplitudes in the field of complex numbers $\C$ by discrete fields
$\ff{p^{2}}$.  If we restrict ourselves to odd primes with
$p\equiv 3\mod 4$, the fields $\ff{p}$ are exactly those for which
$x^2+1$ has no root in the field, and we can adjoin $\ComplexI$ as a
root of $x^2+1$ to the field $\ff{p}$.  For a given element
$\alpha \in \ff{p^2}$, we define the equivalent of complex conjugation
as the map $\alpha\mapsto\alpha^p$, which we may write as
$\alpha^*=\alpha^p$. The resulting extended field is isomorphic to
$\ff{p^2}$ and is a realization of discrete complex numbers suitable
for quantum mechanical calculations.

@@

\mathbf{H}_{1}=\begin{pmatrix}-1 & -\ComplexI & 0\\
\ComplexI & 1 & -\ComplexI\\
0 & \ComplexI & 1
\end{pmatrix}

@@

\mathbf{H}_{2}=\begin{pmatrix}-1 & -1-\ComplexI\\
-1+\ComplexI & 0
\end{pmatrix}

@@

-1+\ComplexI

@@

\mathbf{H}_{3}=\begin{pmatrix}-1 & -1-\ComplexI\\
-1+\ComplexI & 1-
\end{pmatrix}

@@

are $-1-\ComplexI$ with $\begin{pmatrix}1\\
-1-\ComplexI
\end{pmatrix}$ and $-1+\ComplexI$ with $\begin{pmatrix}1\\
1+\ComplexI

@@

-1\pm\sqrt{2} & \equiv & -1\pm\ComplexI\left(\bmod p\right)

@@

For $i=0,1,2$ let $U_i$ be the unitary map that permutes the basis
vectors $\ket{0}$ and $\ket{i}$ and acts as the identity for the rest.
Note that $\ket{\Psi}$ is invariant under $U_i$. So again by property
({\bf iii}) we have
\[
\dpt\big(\ket{\Psi},\ket{i}\big) =\dpt\big(U_i\ket{\Psi},U_i\ket{0}\big)=
\dpt\big(\ket{\Psi},\ket{0}\big) = \frac{1}{2}
\]
for $i=0,1,2$.  These probability assignments are inconsistent with
the requirement that the probabilities for orthogonal outcomes add up
to 1. By property ({\bf i})
\begin{align*}
  1&=\dpt\big(\ket{\Psi},\ket{0}\big)+ \dpt\big(\ket{\Psi},\ket{1}\big) + 
     \dpt\big(\ket{\Psi},\ket{2}\big)+ \dpt \big(\ket{\Psi},\ket{3}\big)\\
  &=\frac{1}{2}+\frac{1}{2}+\frac{1}{2}+\dpt\big(\ket{\Psi},\ket{3}\big)\\
  &\geq \frac{3}{2},
\end{align*}
so we have a contradiction. The assumption of a satisfactory map
$\dpt$ is thus false. $\Box$
\end{proof}
(It should be pointed out that the argument above explicitly uses
states from a vector space of four dimensions or larger, and thus does
not hold if the largest vector space one can consider has dimension
$d\le 3$.  But it does hold for any system that, for example, contains
more than a single qubit.)

@@

  Following the general structure in~\cite{peres1995quantum}, we will
  prove by contradiction that the theorem holds for primes (with
  $p \equiv 3 \mod 4$) with $p > 3$ ~\footnote{\label{fn:p>3}When
    $p=3$, the Kochen-Specker theorem still holds, but the proof is
    longer.}. Suppose $v$ exists, and consider the set of all vectors
  with entries in $\left\{ 0,\pm1,\pm1\pm\ComplexI\right\} $. Then the
  steps in the proof can be read off from Table
  \ref{tab:proof-of-K-S}.  In each cell in Table
  \ref{tab:proof-of-K-S}, the string $xyz$ represents a state vector,
  where $x$, $y$, and $z$ can be \{ $0$, $1$, $\bar{1}\equiv -1,$
  ${+} \equiv (1+\ComplexI)$, ${-} \equiv (1-\ComplexI)$ \}, and
\[
v\left( xyz \right)= \{\textrm{possible}\, \mbox{\it or\/}\,
\textrm{impossible} \ , \]
depending on its column.  Examining the table from top to the bottom,
we will eventually get a contradiction.  The set of vectors are
invariant under interchanges of the $x$, $y$, and $z$ axess.  Hence,
we can assume
$v\left(100\right)=v\left(010\right)=\textrm{impossible}$ and
$v\left(001\right)=\textrm{possible}$.  Furthermore, $110$ and
$1\bar{1}0$ are orthogonal to $001$, so that
$v\left(110\right)=v\left(1\bar{1}0\right)=\textrm{impossible}$.  A
similar argument may be used to construct the second, third, and
fourth rows.  After we know enough values of
$v\left(\ket{\Psi}\right)$, we can find that $10{-}$, ${+}0\bar{1}$,
and $010$ is an orthogonal basis, and we already know that
$v\left({+}0\bar{1}\right)=v\left(010\right)=\textrm{impossible}$, so
$v\left(10{-}\right)=\textrm{possible}$.  Following a similar
procedure, we can continue evaluating the entire content of Table
\ref{tab:proof-of-K-S}.  We can see that $100$, $0{+}1$, and
$0\bar{1}{-}$ form an orthogonal basis, but we already know
$v\left(100\right)=v\left(0{+}1\right)=v\left(0\bar{1}{-}\right)
=\textrm{impossible}$.
This contradicts our initial assumptions. Therefore, there is no
function $v$ satisfying the conditions.
\end{proof}
Notice that this proof applies for almost all $p$, and, despite
the exclusive appearance of integers in $\ff{p^{2}}$ throughout the
computation, there are no occurrences of wrapping back to zero in the
number field.

>>>>

\documentclass[english,12pt]{iopart}
\usepackage{mathrsfs}
\usepackage{fullpage}
\usepackage{url}
\usepackage{times}
\usepackage{amstext}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{iopams}
\usepackage{setstack}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% these went missing?
% \newcommand{\bra}[1]{\left\langle{#1}\right|}
% \newcommand{\ket}[1]{\left|{#1}\right\rangle}
% \newcommand{\mathscr}[1]{{\cal #1}}
%\renewcommand{\author}[1]{\bf #1 \/}
\newcommand{\affiliation}[1]{\parbox{5.5in}{\small{\it #1 \/}}}

@@

\newcommand{\ComplexI}{\text{i}}
\newcommand{\fnorm}[1]{\mathsf{N}(#1)}
\newcommand{\mod}{\mbox{mod }}

@@

@@

Our fundamental divergence from CQT is to replace continuous complex
amplitudes in the field of complex numbers $\C$ by discrete fields
$\ff{p^{2}}$.  If we restrict ourselves to odd primes with
$p\equiv 3\mod 4$, the fields $\ff{p}$ are exactly those for which
$x^2+1$ has no root in the field, and we can adjoin $\rmi$ as a
root of $x^2+1$ to the field $\ff{p}$.  For a given element
$\alpha \in \ff{p^2}$, we define the equivalent of complex conjugation
as the map $\alpha\mapsto\alpha^p$, which we may write as
$\alpha^*=\alpha^p$. The resulting extended field is isomorphic to
$\ff{p^2}$ and is a realization of discrete complex numbers suitable
for quantum mechanical calculations.

@@

\mathbf{H}_{1}=\left(\begin{array}{ccc}
-1 & -\rmi & 0\\
\rmi & 1 & -\rmi\\
0 & \rmi & 1
\end{array}\right)

@@

\mathbf{H}_{2}=\left(\begin{array}{cc}
-1 & -1-\rmi\\
-1+\rmi & 0
\end{array}\right)

@@

-1+\rmi

@@

\mathbf{H}_{3}=\left(\begin{array}{cc}
-1 & -1-\rmi\\
-1+\rmi & 1-
\end{array}\right)

@@

are $-1-\rmi$ with $\begin{pmatrix}1\\
-1-\rmi
\end{pmatrix}$ and $-1+\rmi$ with $\begin{pmatrix}1\\
1+\rmi

@@

-1\pm\sqrt{2} & \equiv & -1\pm\rmi\left(\bmod p\right)

@@

For $i=0,1,2$ let $U_i$ be the unitary map that permutes the basis
vectors $\ket{0}$ and $\ket{i}$ and acts as the identity for the rest.
Note that $\ket{\Psi}$ is invariant under $U_i$. So again by property
({\bf iii}) we have
\[
\dpt\big(\ket{\Psi},\ket{i}\big) =\dpt\big(U_i\ket{\Psi},U_i\ket{0}\big)=
\dpt\big(\ket{\Psi},\ket{0}\big) = \frac{1}{2}
\]
for $i=0,1,2$.  These probability assignments are inconsistent with
the requirement that the probabilities for orthogonal outcomes add up
to 1. By property ({\bf i})
\begin{eqnarray*}
1 & = & \dpt\big(\ket{\Psi},\ket{0}\big)+\dpt\big(\ket{\Psi},\ket{1}\big)+\dpt\big(\ket{\Psi},\ket{2}\big)+\dpt\big(\ket{\Psi},\ket{3}\big)\\
 & = & \frac{1}{2}+\frac{1}{2}+\frac{1}{2}+\dpt\big(\ket{\Psi},\ket{3}\big)\\
 & \geq & \frac{3}{2},
\end{eqnarray*}
so we have a contradiction. The assumption of a satisfactory map
$\dpt$ is thus false. $\Box$
\end{proof}
(It should be pointed out that the argument above explicitly uses
states from a vector space of four dimensions or larger, and thus does
not hold if the largest vector space one can consider has dimension
$d\le 3$.  But it does hold for any system that, for example, contains
more than a single qubit.)

@@

  Following the general structure in~\cite{peres1995quantum}, we will
  prove by contradiction that the theorem holds for primes (with
  $p \equiv 3 \mod 4$) with $p > 3$ ~\footnote{\label{fn:p>3}When
    $p=3$, the Kochen-Specker theorem still holds, but the proof is
    longer.}. Suppose $v$ exists, and consider the set of all vectors
  with entries in $\left\{ 0,\pm1,\pm1\pm\rmi\right\} $. Then the
  steps in the proof can be read off from Table
  \ref{tab:proof-of-K-S}.  In each cell in Table
  \ref{tab:proof-of-K-S}, the string $xyz$ represents a state vector,
  where $x$, $y$, and $z$ can be \{ $0$, $1$, $\bar{1}\equiv -1,$
  ${+} \equiv (1+\rmi)$, ${-} \equiv (1-\rmi)$ \}, and
\[
v\left( xyz \right)= \{\textrm{possible}\, \mbox{\it or\/}\,
\textrm{impossible} \ , \]
depending on its column.  Examining the table from top to the bottom,
we will eventually get a contradiction.  The set of vectors are
invariant under interchanges of the $x$, $y$, and $z$ axess.  Hence,
we can assume
$v\left(100\right)=v\left(010\right)=\textrm{impossible}$ and
$v\left(001\right)=\textrm{possible}$.  Furthermore, $110$ and
$1\bar{1}0$ are orthogonal to $001$, so that
$v\left(110\right)=v\left(1\bar{1}0\right)=\textrm{impossible}$.  A
similar argument may be used to construct the second, third, and
fourth rows.  After we know enough values of
$v\left(\ket{\Psi}\right)$, we can find that $10{-}$, ${+}0\bar{1}$,
and $010$ is an orthogonal basis, and we already know that
$v\left({+}0\bar{1}\right)=v\left(010\right)=\textrm{impossible}$, so
$v\left(10{-}\right)=\textrm{possible}$.  Following a similar
procedure, we can continue evaluating the entire content of Table
\ref{tab:proof-of-K-S}.  We can see that $100$, $0{+}1$, and
$0\bar{1}{-}$ form an orthogonal basis, but we already know
$v\left(100\right)=v\left(0{+}1\right)=v\left(0\bar{1}{-}\right)
=\textrm{impossible}$.
This contradicts our initial assumptions. Therefore, there is no
function $v$ satisfying the conditions.
\end{proof}
Notice that this proof applies for almost all $p$, and, despite
the exclusive appearance of integers in $\ff{p^{2}}$ throughout the
computation, there are no occurrences of wrapping back to zero in the
number field.
